Exploring the World Inside US Prisons for Safety, Health, and Isolation

### 2024-02-15 outputs/paper/paper.qmd
update

## Source
The paper used for replication is Inside the Box: Safety, Health, and Isolation in Prison (@western2021inside), which mainly analyzes the impact of prisons on safety and health, and
specifically includes data from the new coronavirus pandemic.

Sourcebook of Criminal Justice Statistics   
Bureau of Justice Statistics Report      
Jail Inmates in 2018, U.S. Population FRED Data     
World Prison Population Data      
Western B. and Pettit B data      
Survey of Inmates of States Correctional Facilities (1986, 1991, 1997, 2004)    
Survey of Prison Inmates (2016) (Bureau of Justice Statistics 1994, 2004a, 2004b, 2006, 2019, 2021)

## Methodology
All data are not obtained directly through surveys or sampling statistics; rather, they are derived through meticulous processing and calculation based on pre-existing datasets.

@@

## data cleaning and construction
Accurate data cleansing and structuring are paramount in research. Given the extensive and varied raw data analyzed in the original paper Inside the Box: Safety, Health, and Isolation in Prison (@western2021inside) and the lack of detailed explanation regarding data processing methods by the original author, Figures 1, 2, 4, 5, 6, and 7 are derived from processed data provided by the original author, albeit with a distinct focus in visualization. For figure 2, we obtained the updated version of the original data and expanded it by incorporating data from an additional 5 countries to enhance comprehensiveness. @tbl-country

@@

## Measurement
For @fig-over
The Bureau of Justice Statistics (BJS) conducts thorough data collection and analysis on the populations incarcerated in both state and federal prisons, as well as local jails, across the United States. Utilizing diverse sources such as surveys, administrative records, and inmate censuses, BJS produces precise estimates of the total number of individuals held in correctional facilities.

For @fig-admit and @fig-plane,original data comes from two prison questionnaires    
For example, the sample for the 2004 survey was drawn from two different documents. The primary archive includes a list of 1,549 state prisons from the June 30, 2000, BJS 2000 Census of State and Federal Correctional Facilities. The secondary archive includes 36 prisons open between June 30, 2000, and April 1, 2003. The sampling design required a stratified two-stage selection process. In the initial phase, 14 male and 7 female facilities were identified based on the gender ratio within prisons. The remaining facilities are divided into 16 levels based on geographic region and the male and female population within each prison. The weighting procedure involves assigning each inmate a base weight and three adjustment factors to obtain the final weight for the survey. Data was collected through face-to-face interviews with prisoners using computer-assisted personal interviewing techniques.

# Result {#sec-result}
## Over trend

@@

This underscores the continued importance of mass incarceration as a pressing issue in the United States. It highlights the need for further research to fully examine its effects on various aspects including health, property ownership and personal growth. Furthermore, understanding its socioeconomic impact and exploring effective rehabilitation and reintegration strategies are important steps to address this multifaceted problem and build a more just and equitable society.
@fig-over
```{r fig.height=10, fig.width=6 }
#| label: fig-over
#| fig-cap: US Incarceration, 1925-2018
#| echo: false
#| warning: false
#| message: false

@@

## Exploring Racial Disparities in Incarceration Trends
Despite extensive discussions surrounding the surge in the U.S. prison population over the last twenty-five years (@pettit2004mass), scant attention has been paid to the evolution of incarceration inequality. Our study delves into sentencing disparities by analyzing the likelihood of incarceration among blacks and whites across varying educational attainment levels. The data(@fig-men) reveals stark contrasts: irrespective of educational achievement, black individuals face significantly higher incarceration rates compared to their white counterparts, particularly among those with less than twelve years of schooling.

## Global Perspectives on Mass Incarceration
In addition to examining the trajectory of incarceration rates within the United States over time, it is equally imperative to conduct horizontal comparisons by scrutinizing incarceration rates across different countries. Such comparisons enable us to discern whether mass incarceration is a global phenomenon or if it manifests as a more acute issue within the United States.

@@

c_data <- data[order(-data$prison_population_rate), ]

# Create the bar plot
plot <- ggplot(c_data, aes(x =prison_population_rate, y = country)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  theme(axis.text.x = element_text(angle = 0, hjust = 1, vjust = 0.5)) +  # Rotate x-axis labels
  labs(x = "Prison Population Rate", y = "country")  # Add axis labels

@@

# Discussion {#sec-dis}
## What is done in this paper
Through rigorous data collection, cleaning, and analysis procedures, we meticulously obtained vital incarceration rate data spanning from 1925 to 2018 for the United States. Additionally, we meticulously compiled statistics on Men’s cumulative imprisonment risk, organized by cohort, and meticulously gathered data on the incarceration rates of 20 countries worldwide, including the United States, up to 2023.

@@

## rehabilitating
Rehabilitation refers to a series of interventions aimed at enhancing functioning and diminishing disability in individuals with health conditions, taking into account their interactions with the environment.Academics studying mass incarceration identify the 1970s as a critical juncture in U.S. penal history, characterized by a transition towards stricter punitive measures and a widespread belief that traditional rehabilitation methods were ineffective(@Rehabilitation).

@@

```
The original data table exclusively focused on a subset of admitted individuals for the analysis of the American Prison Rehabilitation Program. Consequently, we opted to select the not admitted portion of the dataset to create similar charts for comparative analysis. @fig-plane illustrates that there is minimal disparity between the two groups regarding participation in drug rehabilitation and employment programs. However, the non-admitted group exhibits higher participation rates in education and vocational training initiatives.

@@

## Cumulative COVID-19 Case Rates among Those in Prison and General Population
The closed and densely populated environment within prisons creates ideal conditions for the proliferation of large infectious diseases. The data stemming from the recent COVID-19 pandemic serves as a poignant illustration of this perspective, offering detailed insights into the heightened vulnerability of incarcerated populations to the rapid transmission and severe impact of contagious illnesses.

@@

In addition to replicating the original data, we have incorporated a comparison of the COVID-19 infection rate among prison staff with the overall U.S. COVID-19 infection rate. According to the figure(@fig-staff), the infection rate among prison staff surpasses the overall infection rate but remains lower than the new coronavirus infection rate among incarcerated individuals, positioning it roughly in the middle between the two. This observation underscores the vulnerable position of prisoners within correctional facilities during the pandemic.
```{r fig.height=10, fig.width=10 }
#| label: fig-planes
#| fig-cap: Cumulative COVID-19 Case Rates among Those in Imprisoned persons and General Population, by State
#| echo: false
#| warning: false
#| message: false

@@

## Weaknesses
### Bias in Data Analysis
On one hand, it's noteworthy that a considerable portion of data regarding U.S. prisons exclusively pertains to men, leading to the phenomenon referred to as "big dick data." This term underscores the prevalence of male-dominated datasets, which not only underscores systemic bias and marginalization, particularly towards women but also sheds light on existing power dynamics. The absence or marginalization of women in these datasets not only mirrors prevailing power structures but also perpetuates gender-based biases and stereotypes. Such gender disparity distorts our comprehension of reality, reinforces harmful stereotypes and inequalities, and presents substantial hurdles to data-driven decision-making (@dignazio2020data).

@@

### Measurement error
Utilizing data from various origins, a practice known as data fusion or integration, introduces complexities and hurdles that can result in measurement inaccuracies. Each dataset carries its unique biases, constraints, and inaccuracies stemming from variations in data collection methods, sampling approaches, measurement tools, and processing methodologies.

@@

## future
On the one hand, similar data collection, statistical analysis, and comparative studies of Canadian prisons provide an attractive avenue for gaining insights due to Canada's geographic proximity to the United States and some shared cultural and legal frameworks. This similarity in land area and certain social aspects may allow for more meaningful comparisons and broader conclusions.

>>>>

## Source

The paper used for replication is Inside the Box: Safety, Health, and Isolation in Prison (@western2021inside), which mainly analyzes the impact of prisons on safety and health, and specifically includes data from the new coronavirus pandemic.

Sourcebook of Criminal Justice Statistics\
Bureau of Justice Statistics Report\
Jail Inmates in 2018, U.S. Population FRED Data\
World Prison Population Data\
Western B. and Pettit B data\
Survey of Inmates of States Correctional Facilities (1986, 1991, 1997, 2004)\
Survey of Prison Inmates (2016) (Bureau of Justice Statistics 1994, 2004a, 2004b, 2006, 2019, 2021)

## Methodology

All data are not obtained directly through surveys or sampling statistics; rather, they are derived through meticulous processing and calculation based on pre-existing datasets.

@@

## data cleaning and construction

In research, accurate data cleansing and structuring are imperative. Figures 1, 3, 4, 5, 6, and 7 in this study stem from the original paper Inside the Box: Safety, Health, and Isolation in Prison (@western2021inside). However, due to the original author's lack of detailed explanation regarding data processing methods, we took steps to ensure the reliability and relevance of the data. Specifically, we focused on visualizing the processed data to enhance clarity and understanding.

For Figure 2, we not only obtained the updated version of the original data but also expanded it by including data from an additional 5 countries, as outlined in @tbl-country. This expansion was undertaken to augment the comprehensiveness of our analysis. However, to maintain the relevance and accuracy of the displayed data, we made a deliberate decision not to include countries with missing data sources from the website.

@@

## Measurement

For @fig-over The Bureau of Justice Statistics (BJS) conducts thorough data collection and analysis on the populations incarcerated in both state and federal prisons, as well as local jails, across the United States. Utilizing diverse sources such as surveys, administrative records, and inmate censuses, BJS produces precise estimates of the total number of individuals held in correctional facilities.

For @fig-admit and @fig-plane,original data comes from two prison questionnaires\
For example, the sample for the 2004 survey was drawn from two different documents. The primary archive includes a list of 1,549 state prisons from the June 30, 2000, BJS 2000 Census of State and Federal Correctional Facilities. The secondary archive includes 36 prisons open between June 30, 2000, and April 1, 2003. The sampling design required a stratified two-stage selection process. In the initial phase, 14 male and 7 female facilities were identified based on the gender ratio within prisons. The remaining facilities are divided into 16 levels based on geographic region and the male and female population within each prison. The weighting procedure involves assigning each inmate a base weight and three adjustment factors to obtain the final weight for the survey. Data was collected through face-to-face interviews with prisoners using computer-assisted personal interviewing techniques.

# Result {#sec-result}

## Over trend

@@

This underscores the continued importance of mass incarceration as a pressing issue in the United States. It highlights the need for further research to fully examine its effects on various aspects including health, property ownership and personal growth. Furthermore, understanding its socioeconomic impact and exploring effective rehabilitation and reintegration strategies are important steps to address this multifaceted problem and build a more just and equitable society. @fig-over

```{r fig.height=10, fig.width=6 }
#| label: fig-over
#| fig-cap: US Incarceration, 1925-2018
#| echo: false
#| warning: false
#| message: false

@@

## Exploring Racial Disparities in Incarceration Trends

Despite extensive discussions surrounding the surge in the U.S. prison population over the last twenty-five years (@pettit2004mass), scant attention has been paid to the evolution of incarceration inequality. Our study delves into sentencing disparities by analyzing the likelihood of incarceration among blacks and whites across varying educational attainment levels. The data(@fig-men) reveals stark contrasts: irrespective of educational achievement, black individuals face significantly higher incarceration rates compared to their white counterparts, particularly among those with less than twelve years of schooling.

## Global Perspectives on Mass Incarceration

In addition to examining the trajectory of incarceration rates within the United States over time, it is equally imperative to conduct horizontal comparisons by scrutinizing incarceration rates across different countries. Such comparisons enable us to discern whether mass incarceration is a global phenomenon or if it manifests as a more acute issue within the United States.

@@

c_data <- data[order(-data$prison_rate_per100000), ]

# Create the bar plot
plot <- ggplot(c_data, aes(x =prison_rate_per100000, y = country)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  theme(axis.text.x = element_text(angle = 0, hjust = 1, vjust = 0.5)) +  # Rotate x-axis labels
  labs(x = "Prison Rate (per 100,000)", y = "country")  # Add axis labels

@@

# Discussion {#sec-dis}

## What is done in this paper

Through rigorous data collection, cleaning, and analysis procedures, we meticulously obtained vital incarceration rate data spanning from 1925 to 2018 for the United States. Additionally, we meticulously compiled statistics on Men’s cumulative imprisonment risk, organized by cohort, and meticulously gathered data on the incarceration rates of 20 countries worldwide, including the United States, up to 2023.

@@

## rehabilitating

Rehabilitation refers to a series of interventions aimed at enhancing functioning and diminishing disability in individuals with health conditions, taking into account their interactions with the environment.Academics studying mass incarceration identify the 1970s as a critical juncture in U.S. penal history, characterized by a transition towards stricter punitive measures and a widespread belief that traditional rehabilitation methods were ineffective(@Rehabilitation).

@@

The original data table exclusively focused on a subset of admitted individuals for the analysis of the American Prison Rehabilitation Program. Consequently, we opted to select the not admitted portion of the dataset to create similar charts for comparative analysis. @fig-plane illustrates that there is minimal disparity between the two groups regarding participation in drug rehabilitation and employment programs. However, the non-admitted group exhibits higher participation rates in education and vocational training initiatives.

@@

## Cumulative COVID-19 Case Rates among Those in Prison and General Population

The closed and densely populated environment within prisons creates ideal conditions for the proliferation of large infectious diseases. The data stemming from the recent COVID-19 pandemic serves as a poignant illustration of this perspective, offering detailed insights into the heightened vulnerability of incarcerated populations to the rapid transmission and severe impact of contagious illnesses.

@@

In addition to replicating the original data, we have incorporated a comparison of the COVID-19 infection rate among prison staff with the overall U.S. COVID-19 infection rate. According to the figure(@fig-staff), the infection rate among prison staff surpasses the overall infection rate but remains lower than the new coronavirus infection rate among incarcerated individuals, positioning it roughly in the middle between the two. This observation underscores the vulnerable position of prisoners within correctional facilities during the pandemic.

```{r fig.height=10, fig.width=10 }
#| label: fig-planes
#| fig-cap: Cumulative COVID-19 Case Rates among Those in Imprisoned persons and General Population, by State
#| echo: false
#| warning: false
#| message: false

@@

## Weaknesses

### Bias in Data Analysis

On one hand, it's noteworthy that a considerable portion of data regarding U.S. prisons exclusively pertains to men, leading to the phenomenon referred to as "big dick data." This term underscores the prevalence of male-dominated datasets, which not only underscores systemic bias and marginalization, particularly towards women but also sheds light on existing power dynamics. The absence or marginalization of women in these datasets not only mirrors prevailing power structures but also perpetuates gender-based biases and stereotypes. Such gender disparity distorts our comprehension of reality, reinforces harmful stereotypes and inequalities, and presents substantial hurdles to data-driven decision-making (@dignazio2020data).

@@

### Measurement error

Utilizing data from various origins, a practice known as data fusion or integration, introduces complexities and hurdles that can result in measurement inaccuracies. Each dataset carries its unique biases, constraints, and inaccuracies stemming from variations in data collection methods, sampling approaches, measurement tools, and processing methodologies.

@@

## future

On the one hand, similar data collection, statistical analysis, and comparative studies of Canadian prisons provide an attractive avenue for gaining insights due to Canada's geographic proximity to the United States and some shared cultural and legal frameworks. This similarity in land area and certain social aspects may allow for more meaningful comparisons and broader conclusions.
